---
title: "Sea cucumbers may diversify shellfish aquaculture"
author: "Frank DAVID"
date: '2023-01-30'
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME", "C")
```

### R Markdown

This R Markdown document is made to support the findings of the paper "Sea cucumber juveniles show great potential to diversify shellfish aquaculture in temperate zones" by David et al. (2023). 


## Preparing the toolbox

Let's begin by loading the libraries that we will require to run the analyses.

```{r echo=TRUE, message=FALSE}
library(agricolae)
library(vegan)
library(ade4)
library(factoextra)
```


## Loading the data
We load growth, pigments, amino acid, fatty acid and stable isotope data from the GitHub server.
For details on line and column headings please check the readme file.

```{r}

growth <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-juveniles/main/GrowthHoloFarm.csv", sep=",", header=T)
AA <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-juveniles/main/AAHoloFarm.csv", sep=",", header=T)
FA <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-juveniles/main/FAHoloFarm.csv", sep=",", header=T)
pig <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-juveniles/main/PigHoloFarm.csv", sep=",", header=T)
SI <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-juveniles/main/SIHoloFarm.csv", sep=",", header=T)

```




# Growth and survival

```{r}
# Mean weight and standard deviation

tapply(growth$Idweight, paste(growth$Origin, growth$Time), mean)
tapply(growth$Idweight, paste(growth$Origin, growth$Time), sd)

#Specific growth rates calculation

SGRplo1=log(mean(growth$Idweight[growth$Origin=="Plouguerneau" & growth$Time=="J133"])/mean(growth$Idweight[growth$Origin=="Plouguerneau" & growth$Time=="J0"]))*100/133

SGRplo2=(log(mean(growth$Idweight[growth$Origin=="Plouguerneau" & growth$Time=="J281"]))-log(mean(growth$Idweight[growth$Origin=="Plouguerneau" & growth$Time=="J133"])))*100/(281-133)

SGRbre=(log(mean(growth$Idweight[growth$Origin=="Brest" & growth$Time=="J178"]))-log(mean(growth$Idweight[growth$Origin=="Brest" & growth$Time=="J0"])))*100/178

SGRcan=(log(mean(growth$Idweight[growth$Origin=="Cancale" & growth$Time=="J133"]))-log(mean(growth$Idweight[growth$Origin=="Cancale" & growth$Time=="J0"])))*100/133

SGRAqB=(log(mean(growth$Idweight[growth$Origin=="AQUA-B" & growth$Time=="J125"]))-log(mean(growth$Idweight[growth$Origin=="AQUA-B" & growth$Time=="J0"])))*100/125

```

## Graphical representation

```{r echo=FALSE, fig.width=3.5, fig.height=7, fig.align="center"}
  
initial=as.POSIXct("01/10/20", format = "%d/%m/%y")
spl=as.POSIXct(growth$Date, format = "%d/%m/%y")
lvlspl=as.POSIXct(levels(as.factor(spl)), format = "%Y-%m-%d")
  
#svg(file="fig2.svg", width=3.5, height=7)
  
  par(las=1, mar=c(4,5,1,2), mfrow=c(4,1), bty="n", oma=c(0,0,1,0))
  
  plot(1, 1, type="n", xlim=c(initial, max(lvlspl)), ylim=c(0, 60), ylab="Individuals weight (g)", xaxt="n", xlab="")
  axis.POSIXct(1, at=c(initial, lvlspl[c(1, 3, 6)]), format ="%d/%m/%y", labels=T, las=1)
  text(initial +3000000, 50, "Survival : 93% \nInitial n = 80", cex=0.8, adj=c(0,1))
  text(mean(lvlspl[c(1, 3)]), 25, bquote("SGR ="~ "0.28 %" ~ d^-1), cex=0.7, adj=0.5)
  text(mean(lvlspl[c(3, 6)]), 40, bquote("SGR ="~ "0.88 %" ~ d^-1), cex=0.7, adj=0.5)
  boxplot(growth$Idweight[growth$Origin=="Plouguerneau"]~spl[growth$Origin=="Plouguerneau"], at=lvlspl[c(1, 3, 6)], add=T, boxwex=2000000, ann=F, xaxt="n", yaxt="n")
  title(main="Sea ranched with European abalones (Ea)", cex.main=0.9)
    
  plot(1, 1, type="n", xlim=c(initial, max(lvlspl)), ylim=c(0, 60), ylab="Individuals weight (g)", xaxt="n", xlab="")
  axis.POSIXct(1, labels=T, at=c(initial, lvlspl[c(2,9)]), format ="%d/%m/%y", las=1)
  text(initial +3000000, 50, "Survival : 100% \nInitial n = 18", cex=0.8, adj=c(0,1))
  text(mean(lvlspl[c(2, 9)]), 40, bquote("SGR ="~ "0.94 %" ~ d^-1), cex=0.7, adj=0.5)
  boxplot(growth$Idweight[growth$Origin=="Brest"]~spl[growth$Origin=="Brest"], at=lvlspl[c(2, 9)], add=T, boxwex=2000000, ann=F, xaxt="n", yaxt="n")
  title(main="Sea ranched underneath European flat oysters (Eo)", cex.main=0.9)
  
  plot(1, 1, type="n", xlim=c(initial, max(lvlspl)), ylim=c(0, 60), ylab="Individuals weight (g)", xaxt="n", xlab="")
  axis.POSIXct(1, at=c(initial, lvlspl[c(4,8)]), format ="%d/%m/%y", labels=T, las=1)
  text(initial +3000000, 50, "Survival : 99% \nInitial n = 81", cex=0.8, adj=c(0,1))
  text(mean(lvlspl[c(4, 8)]), 25, bquote("SGR ="~ "0.24 %" ~ d^-1), cex=0.7, adj=0.5)
  boxplot(growth$Idweight[growth$Origin=="Cancale"]~spl[growth$Origin=="Cancale"], at=lvlspl[c(4, 8)], add=T, boxwex=2000000, ann=F, xaxt="n", yaxt="n")
  title(main="Sea ranched underneath Pacific oysters (Po)", cex.main=0.9)
  
  plot(1, 1, type="n", xlim=c(initial, max(lvlspl)), ylim=c(0, 60), ylab="Individuals weight (g)", xaxt="n", xlab="")
  axis.POSIXct(1, at=c(initial, lvlspl[c(5,7)]), format ="%d/%m/%y", labels=T, las=1)
  text(initial +3000000, 50, "Survival : 84% \nInitial n = 240", cex=0.8, adj=c(0,1))
  text(mean(lvlspl[c(5, 7)]), 25, bquote("SGR ="~ "2.76 %" ~ d^-1), cex=0.7, adj=0.5)
  boxplot(growth$Idweight[growth$Origin=="AQUA-B"]~spl[growth$Origin=="AQUA-B"], at=lvlspl[c(5, 7)], add=T, boxwex=2000000, ann=F, xaxt="n", yaxt="n")
  title(main="Indoor monoculture (Im)", cex.main=0.9)
  
#dev.off()

```



# Pigments
##Total pigments description
```{r}

# Concentration of pigments per mg of dry stomach content

pigconc=pig[,-c(1:4)]*pig$solventVol/pig$splMass
rownames(pigconc)=pig[,1]

# Mean total sum of pigments and standard deviation

totpig=rowSums(pigconc)
tapply(totpig, pig$condition, mean)
tapply(totpig, pig$condition, sd)

# Statistical comparisons between treatments

waerden.test(totpig[-c(1:2)], pig$condition[-c(1:2)], console=T)
```
## Total pigments visualisation
```{r echo=FALSE, fig.width=6, fig.height=4, fig.align="center"}

#pdf(file="fig3.pdf", width=5, height=4)

  par(las=1, mar=c(3,5,1,2), bty="n", mfrow=c(1,1))
  boxplot(totpig~pig$condition, xlab="", ylab=expression(paste("Pigment concentration (mg g d", w^-1, ")")), names=c("Ea", "Eo", "Po", "Im","Im feed"), ylim=c(0, 110))
  text(1:5, tapply(totpig, pig$condition, max)+8, labels=c("b", "c", "d", "a", ""), font=1, cex=1.5)

#dev.off()
  
```


## Summary of pigments table
```{r}
# Let's group main pigments and their derivatives

pigconc2=cbind(pigconc[,2], rowSums(pigconc[,c(1,3,4)]), pigconc[,c(5:7)], rowSums(pigconc[,c(8:9)]), pigconc[,c(10:11)], rowSums(pigconc[,c(12:16)]), rowSums(pigconc[,c(17:20)]), rowSums(pigconc[,c(21:23)]), rowSums(pigconc[,c(24:25)]))
names(pigconc2)=c("Fuco", "Fuco like", "Diadino", "Diato", "Lutein", "Zeax", "Cantha", "Chloro-b", "B-car", "Pheopb-a", "Chloro-a", "Pheopt-a")

# Let's edit a summary table

summaryPIG=data.frame(M=aggregate(cbind(pigconc2, totpig), by=list(pig$condition), mean), SD=aggregate(cbind(pigconc2, totpig), by=list(pig$condition), sd), LENGTH=aggregate(cbind(pigconc2, totpig), by=list(pig$condition), length))
```

## Principal component analysis on pigment compositions of stomach contents

```{r, echo=FALSE, fig.width=9, fig.height=6, fig.align="center"}

acp=dudi.pca(pigconc2[-c(1:2),], scannf = F, nf = 2, center=T, scale=T)

#svg(file="fig4.svg", width=7.5, height=6)

  par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

  plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, xlim=c(-5.5,5.5), ylim=c(-5,5))
  axis(1, pos=0, at=c(seq(-5, 5)[-6]), labels=c(seq(-5, 5)[-6]), tck=0.01, padj=-4)
  axis(2, pos=0, at=c(seq(-5, 5)[-6]), labels=c(seq(-5, 5)[-6]), tck=0.01, hadj=-0.1)

  title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
  title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)
  
  arrows(x0=0, y0=0, x1=acp$co[,1]*5, y1=acp$co[,2]*5, length=0.1, col="grey40", lty=5)
  text(acp$co*5.4, labels=colnames(pigconc2), col="grey40")
  

  a=ordihull(acp$li, groups=pig$condition[-c(1:2)], lty=4, label=F, display=c("sites"))
  points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey80", "grey80", "grey80", "grey20")[as.factor(pig$condition[-c(1:2)])], cex=1.5)

  centroid=t(summary(a))
  labs=substring(rownames(centroid), 3, nchar(rownames(centroid)))

  rect(xright=centroid[,1]+0.3, ybottom=centroid[,2]-0.4, xleft=centroid[,1]-0.3, ytop=centroid[,2]+0.4, col=adjustcolor("white", alpha=0.8), border="grey")
  text(centroid, labels=labs, font=2)
  
  legend("topleft", legend=c("Sea ranched", "Indoor farmed"), pch=23, col="grey40", text.col="black", pt.bg=c("grey80", "grey20"), bty="n", pt.cex=2, inset=c(0.02, 0))
  legend("bottomright", legend=paste("n =", nrow(acp$li)), bty="n", pt.cex=2)

#dev.off()
  
```

Note that the figure was further modified with a vectorial graphics editor to avoid labels superposition

# Total fatty acids and total amino acids
## Means and standard deviations

```{r}
  
# Let's define grouping factors

AAcond=as.factor(substring(AA[,1], 8, nchar(AA[,1])))
AAcond=factor(AAcond, levels=c("Ea","Eo","Po","Im", "Wild"))
FAcond=as.factor(substring(FA[,1], 8, nchar(FA[,1])))
FAcond=factor(FAcond, levels=c("Ea","Eo","Po","Im", "Wild"))
  
# Absolute fatty acid concentrations

FAconc=FA[,-c(1:3, which(colnames(FA)=="C23.0"))]/FA[,which(colnames(FA)=="C23.0")]*FA[,which(colnames(FA)=="StdC23")]/FA[,which(colnames(FA)=="splMass")]
totFA=rowSums(FAconc)
  
rownames(FAconc)=FA[,1]

# Let's evidence statistical differences on total AA and total FA

waerden.test(AA$PrProt, AAcond, console=T)
waerden.test(totFA, FAcond, console=T)

```


## Graphical representation
```{r echo=FALSE, fig.width=5, fig.height=7, fig.align="center"}

#pdf(file="fig5.pdf", width=5, height=7)

  par(las=1, mar=c(4,5,1,2), mfrow=c(2,1), bty="n", oma=c(0,0,1,0))
  
  boxplot(AA$PrProt*10~AAcond, xlab="", ylab=expression(paste("Amino acids concentration (mg g d", w^-1, ")")), ylim=c(0,600), cex.lab=0.9) 
  text(1:5, tapply(AA$PrProt*10, AAcond, max)+50, labels=c("b", "b", "c", "b", "a"), font=1, cex=1.5)

  boxplot(totFA~FAcond, xlab="", ylab=expression(paste("Fatty acids concentration (mg g d", w^-1, ")")), ylim=c(0,17), cex.lab=0.9)
  text(1:5, tapply(totFA, FAcond, max)+1.5, labels=c("a", "ab", "c", "bc", "d"), font=1, cex=1.5)

#dev.off()

```

# Amino acids
## Summary table

```{r}

#Let's group essential amino acids

EAA=rowSums(AA[, c(which(colnames(AA)=="His"), which(colnames(AA)=="Ile"), which(colnames(AA)=="Leu"), which(colnames(AA)=="Lys"), which(colnames(AA)=="Met"), which(colnames(AA)=="Phe"), which(colnames(AA)=="Thr"), which(colnames(AA)=="Val"))])

#Let's edit a summary table

summaryAA=data.frame(M=aggregate(cbind(AA[,-c(1:3)], EAA, AA$PrProt), by=list(AAcond), mean), SD=aggregate(cbind(AA[,-c(1:3)], EAA, AA$PrProt), by=list(AAcond), sd), LENGTH=aggregate(cbind(AA[,-c(1:3)], EAA, AA$PrProt), by=list(AAcond), length))
```

## Principal component analysis on amino acid compositions of body wall

```{r, echo=FALSE, fig.width=9, fig.height=6, fig.align="center"}

  acp=dudi.pca(AA[,-c(1:3)], scannf = F, nf = 2, center=T, scale=T)
  varPCA=get_pca_var(acp)
  
 
#svg(file="fig6b.svg", width=7.5, height=6)

  par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

  plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, xlim=c(-8,8), ylim=c(-6,6))
  axis(1, pos=0, at=c(seq(-8, 8)[-9]), labels=c(seq(-8, 8)[-9]), tck=0.01, padj=-4)
  axis(2, pos=0, at=c(seq(-6, 5)[-7]), labels=c(seq(-6, 5)[-7]), tck=0.01, hadj=-0.1)

  title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
  title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)
    
  arrows(x0=0, y0=0, x1=acp$co[,1]*6, y1=acp$co[,2]*6, length=0.1, col="grey40", lty=5)
  text(acp$co*6.4, labels=colnames(AA[,-c(1:3)]), col="grey40")
  
  a=ordihull(acp$li, groups=AAcond, lty=4, label=F, display=c("sites"))
  points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey80", "grey80", "grey80", "grey20", "white")[AAcond], cex=1.5)

  centroid=t(summary(a))
  labs=rownames(centroid)

  rect(xright=centroid[,1]+0.5, ybottom=centroid[,2]-0.4, xleft=centroid[,1]-0.5, ytop=centroid[,2]+0.4, col=adjustcolor("white", alpha=0.8), border="grey")
  text(centroid, labels=labs, font=2)
  

  legend("topleft", legend=c("Sea ranched", "Indoor farmed", "Wild specimen"), pch=23, col="grey40", text.col="black", pt.bg=c("grey80", "grey20", "white"), bty="n", pt.cex=2, inset=c(0.02, 0))
  legend("bottomright", legend=paste("n =", nrow(acp$li)), bty="n", pt.cex=2)

#dev.off()

```

Note that the figure was further modified with a vectorial graphics editor to avoid labels superposition


# Fatty acids
## Summary Table
 
```{r}

#Let's convert absolute concentrations to relative abundances

FAprct=FAconc/rowSums(FAconc)*100

#Let's define categories of FA
  
brFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C15.0iso"), which(colnames(FAprct)=="C15.0anteiso"), which(colnames(FAprct)=="C16.0iso"))])
SFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C14.0"), which(colnames(FAprct)=="C15.0"), which(colnames(FAprct)=="C16.0"), which(colnames(FAprct)=="C18.0"), which(colnames(FAprct)=="C19.0"),which(colnames(FAprct)=="C20.0"), which(colnames(FAprct)=="C21.0"), which(colnames(FAprct)=="C22.0"), which(colnames(FAprct)=="C24.0"))])
PUFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C18.2w6"), which(colnames(FAprct)=="C18.3w3"), which(colnames(FAprct)=="C18.4w3"), which(colnames(FAprct)=="C20.2w6"), which(colnames(FAprct)=="C20.4w6"), which(colnames(FAprct)=="C20.5w3"), which(colnames(FAprct)=="C22.5w6"), which(colnames(FAprct)=="C22.6w3"))])
HUFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C20.4w6"), which(colnames(FAprct)=="C20.5w3"), which(colnames(FAprct)=="C22.5w6"), which(colnames(FAprct)=="C22.6w3"))])
w3=rowSums(FAprct[, c(which(colnames(FAprct)=="C18.3w3"), which(colnames(FAprct)=="C18.4w3"), which(colnames(FAprct)=="C20.5w3"), which(colnames(FAprct)=="C22.6w3"))])
w6=rowSums(FAprct[, c(which(colnames(FAprct)=="C18.2w6"), which(colnames(FAprct)=="C20.2w6"), which(colnames(FAprct)=="C20.4w6"), which(colnames(FAprct)=="C22.5w6"))])
LCMUFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C20.1w9"), which(colnames(FAprct)=="C20.1w7"), which(colnames(FAprct)=="C22.1w11"), which(colnames(FAprct)=="C22.1w9"), which(colnames(FAprct)=="C23.1w9"), which(colnames(FAprct)=="C24.1w9"))])
MUFA=rowSums(FAprct[, c(which(colnames(FAprct)=="C16.1w7"), which(colnames(FAprct)=="C18.1w9"), which(colnames(FAprct)=="C18.1w7"), which(colnames(FAprct)=="C20.1w9"), which(colnames(FAprct)=="C20.1w7"), which(colnames(FAprct)=="C22.1w11"), which(colnames(FAprct)=="C22.1w9"), which(colnames(FAprct)=="C23.1w9"), which(colnames(FAprct)=="C24.1w9"))])

# Let's edit a summary table

summaryFA=data.frame(M=aggregate(cbind(FAprct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, totFA) , by=list(FAcond), mean), SD=aggregate(cbind(FAprct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, totFA), by=list(FAcond),sd), LE=aggregate(cbind(FAprct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, totFA), by=list(FAcond),length))
```


## Principal component analysis on fatty acid compositions of body wall

```{r, echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

   acp=dudi.pca(FAprct, scannf = F, nf = 2, center=T, scale=T)
  varPCA=get_pca_var(acp)
  
#svg(file="fig7.svg", width=7.5, height=6)

  par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

  plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, xlim=c(-7,7), ylim=c(-6,6))
  axis(1, pos=0, at=c(seq(-7, 7)[-8]), labels=c(seq(-7, 7)[-8]), tck=0.01, padj=-4)
  axis(2, pos=0, at=c(seq(-6, 6)[-7]), labels=c(seq(-6, 6)[-7]), tck=0.01, hadj=-0.1)

  title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
  title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)
  
  arrows(x0=0, y0=0, x1=acp$co[,1][which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)]*6, y1=acp$co[,2][which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)]*6, length=0.1, col="grey40", lty=5)
  arrows(x0=0, y0=0, x1=acp$co[,1][which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)]*6, y1=acp$co[,2][which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)]*6, length=0.1, col="grey40", lty=5)

  text(acp$co[,1][which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)]*6.8, acp$co[,2][which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)]*6.8, labels=substring(names(which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)), 2, nchar(names(which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*2)))), col="grey40")
  text(acp$co[,1][which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)]*6.2, acp$co[,2][which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)]*6.2, labels=substring(names(which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)), 2, nchar(names(which(varPCA$contrib[,2]>100/nrow(varPCA$contrib)*2)))), col="grey40")

  a=ordihull(acp$li, groups=FAcond, lty=4, label=F, display=c("sites"))
  points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey80", "grey80", "grey80", "grey20", "white")[FAcond], cex=1.5)

  centroid=t(summary(a))
  labs=rownames(centroid)

  rect(xright=centroid[,1]+0.5, ybottom=centroid[,2]-0.4, xleft=centroid[,1]-0.5, ytop=centroid[,2]+0.4, col=adjustcolor("white", alpha=0.8), border="grey")
  text(centroid, labels=labs, font=2)
  

  legend("topleft", legend=c("Sea ranched", "Indoor farmed", "Wild grown"), pch=23, col="grey40", text.col="black", pt.bg=c("grey80", "grey20", "white"), bty="n", pt.cex=2, inset=c(0.02, 0))
  legend("bottomright", legend=paste("n =", nrow(acp$li)), bty="n", pt.cex=2)

#dev.off()


```

Note that the figure was further modified with a vectorial graphics editor to avoid labels superposition


